Anomaly detection with kernel density estimation on manifolds

### 2022-05-22 paper/paper.Rmd
Format

# Acknowledgment {.unnumbered}

@@

The tangent space $T_pM$ also admits a dual space $T^*_pM$ called the
*cotangent space* with corresponding *cotangent vector*
$z_p: T^*_pM \rightarrow \mathbb{R}^d$, and basis denoted as
$dx_1(p),\dots,dx_d(p)$.

@@

The Riemannian metric $g$ can be used to define the norm of a vector
$u$, $\|u\| = \sqrt{<u,v>_g}$ and the angle between two vectors $u$ and
$v$, $\cos\theta = \frac{<u,v>_g}{\|u\| \|v\|}$, which are the geometric
quantities induced by $g$. $g$ could also be used to define the line
element $dl^2 = g_{ij}dx_idx_j$ and the volume element
$dV_g = \sqrt{\det(g)}dx_1\dots dx_d$, where $(x_1,\dots,x_d)$ are the
local coordinates of the chart $(U, \varphi)$. For a curve
$\gamma: I \rightarrow M$, the length of the curve is
$$l(\gamma) = \sqrt{\int_0^1 \|\gamma\prime(t)\|^2_g dt} = \sqrt{\int_0^1 g_{ij} \frac{dx_i}{dt} \frac{dx_j}{dt} dt}$$,
where $\gamma(I) \subset U$. The volume of $W \subset U$ is defined as
$$
Vol(W) = \int_W \sqrt{\det(g)}dx_1 \dots dx_d,
$$ which is also called the *Riemannian measure* on $M$.

@@

Pushforward and pullback are two notions corresponding to the notions of
tangent and cotangent vectors. Let $f$ be an embedding from the
Riemannian manifold $(M,g)$ to another smooth manifold $E$. The
pushforward $h=\varphi*g$ of the Riemannian metric $g$ along
$\varphi \equiv f^{-1}$ is a linear map $f_*:TM \rightarrow TE$, which
maps the tangent vectors $u_p$ at point $p \in M$ to the tangent vectors
$f_*u_p$ at the mapping point $f(p) \in E$, and satisfies that for
$u, v \in T_f(p)E$, $$
<u,v>_{\varphi*g_p} = <df_p^{-1}u, df_p^{-1}v>_{g_p},
$$ where $df_p^{-1}$ is the Jacobian of $f^{-1}$. The tangent vectors
$f_*u_p$ are equivalent to the velocity vector of a curve
$\gamma: I\rightarrow M$ passing through point $p$ at time zero with a
constant speed $\gamma^{-1}(0)=u_p$. Similarly, the pullback maps the
cotangent vectors $z_{f(p)}$ at $f(p) \in E$ to cotangent vectors at
$p \in M$ acting on tangent vectors $u \in T_pM$.

>>>>

# Acknowledgment {-}

@@

The tangent space $T_pM$ also admits a dual space $T^\star_pM$ called the *cotangent space* with corresponding *cotangent vector*
$z_p: T^\star_pM \rightarrow \mathbb{R}^d$, and basis denoted as $dx_1(p),\dots,dx_d(p)$.

@@

The Riemannian metric $g$ can be used to define the norm of a vector $u$, $\|u\| = \sqrt{<u,v>_g}$ and the angle between two vectors $u$ and $v$, $\cos\theta = \frac{<u,v>_g}{\|u\| \|v\|}$, which are the geometric quantities induced by $g$. $g$ could also be used to define the line element $dl^2 = g_{ij}dx_idx_j$ and the volume element $dV_g = \sqrt{\det(g)}dx_1\dots dx_d$, where $(x_1,\dots,x_d)$ are the local coordinates of the chart $(U, \varphi)$. 
For a curve $\gamma: I \rightarrow M$, the length of the curve is
$$
l(\gamma) = \sqrt{\int_0^1 \|\gamma\prime(t)\|^2_g dt} = \sqrt{\int_0^1 g_{ij} \frac{dx_i}{dt} \frac{dx_j}{dt} dt},
$$
where $\gamma(I) \subset U$. The volume of $W \subset U$ is defined as
$$
Vol(W) = \int_W \sqrt{\det(g)}dx_1 \dots dx_d,
$$
which is also called the *Riemannian measure* on $M$.

@@

Pushforward and pullback are two notions corresponding to the notions of tangent and cotangent vectors. Let $f$ be an embedding from the Riemannian manifold $(M,g)$ to another smooth manifold $E$. The pushforward $h=\varphi*g$ of the Riemannian metric $g$ along $\varphi \equiv f^{-1}$ is a linear map $f_*:TM \rightarrow TE$, which maps the tangent vectors $u_p$ at point $p \in M$ to the tangent vectors $f_*u_p$ at the mapping point $f(p) \in E$, and satisfies that for $u, v \in T_f(p)E$, 
$$
<u,v>_{\varphi*g_p} = <df_p^{-1}u, df_p^{-1}v>_{g_p},
$$ 
where $df_p^{-1}$ is the Jacobian of $f^{-1}$. The tangent vectors $f_\star u_p$ are equivalent to the velocity vector of a curve $\gamma: I\rightarrow M$ passing through point $p$ at time zero with a constant speed $\gamma^{-1}(0)=u_p$. Similarly, the pullback maps the cotangent vectors $z_{f(p)}$ at $f(p) \in E$ to cotangent vectors at $p \in M$ acting on tangent vectors $u \in T_pM$.
